---
title: "Stratified LD score regression"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    number_sections: no
    toc: yes
    toc_float: true
editor_options:
  chunk_output_type: console
---

### Prepare input
* PGC3 summary statistics was merged with HapMap3 snplist. 
* Full baseline model (53 annotation) from Alke's group was used together
* LD score was computed using 1000G panel and the same list of SNPs as the baseline's

### Enrichment analysis for pseudo bulk peak sets
The annotation includes differential peaks (1v0hr and 6v0hr bothside of fold change), resting condition peaks called directly through MACS2 by merging cells within the same cell type and time point, and "static" peaks that remains in resting condition after removing the differential peaks.<\br>
Please note that enrichment estimate in S-LDSC is the proportion of heritability ($h^2$) devided by the proportion of causal SNPs. In "Comparison of heritability", we only showed $h^2$.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
#library(rtracklayer)

plot_enrichment = function(dat, title, xlab="Odds ratio", ylab="Annotations") {
  p1 <- ggplot(dat, aes(y=annotation)) + 
    geom_point(aes(x=Enrichment), colour="#800000") + #plot OR
    geom_segment(aes(x=lower,xend=upper,yend=annotation), alpha=0.8)
  p1 <- p1 + geom_vline(xintercept=0) + theme_bw() + labs(title=title, x=xlab, y=ylab)
  p1
}


plot_h2 = function(dat, title, xlab="Odds ratio", ylab="Annotations") {
  p1 <- ggplot(dat, aes(y=annotation)) + 
    geom_point(aes(x=Prop._h2), colour="#800000") + #plot OR
    geom_segment(aes(x=lower,xend=upper,yend=annotation), alpha=0.8)
  p1 <- p1 + geom_vline(xintercept=0) + theme_bw() + labs(title=title, x=xlab, y=ylab)
  p1
}
```

#### Dynamic elements vs. static elements from pseudo bulk (joint analysis)
```{r, echo=F}
enrich.ld <- read.table("data/enrich_results/pseudobulk_dyn_vs_stat_joint.results", sep="\t", header=T, row.names=1)
temp <- enrich.ld[(nrow(enrich.ld)-8):nrow(enrich.ld),4:5]
annot <- c('1vs0_NEFMp',
 '6vs0_GABA',
 '1vs0_GABA',
 '6vs0_NEFMp',
 'NpCm_glut_static',
 '6vs0_NEFMm',
 'NmCp_glut_static',
 'GABA_static',
 '1vs0_NEFMm')
rownames(temp) <- annot
temp$annotation <- factor(rownames(temp), level=rownames(temp)[c(8,7,5,3,2,9,6,1,4)])
temp$lower <- temp$Enrichment - 1.96 * temp$Enrichment_std_error
temp$upper <- temp$Enrichment + 1.96 * temp$Enrichment_std_error
plot_enrichment(temp, "", xlab = 'enrichment of SZ variants', ylab = "Functional annotations")
```

#### Comparison of heritability in pseudo bulk (joint analysis)
```{r, echo=F}
enrich.ld <- read.table("data/enrich_results/pseudobulk_dyn_vs_stat_joint.results", sep="\t", header=T, row.names=1)
temp <- enrich.ld[(nrow(enrich.ld)-8):nrow(enrich.ld),]
annot <- c('1vs0_NEFMp',
 '6vs0_GABA',
 '1vs0_GABA',
 '6vs0_NEFMp',
 'NpCm_glut_static',
 '6vs0_NEFMm',
 'NmCp_glut_static',
 'GABA_static',
 '1vs0_NEFMm')
rownames(temp) <- annot
temp$annotation <- factor(rownames(temp), level=rownames(temp)[c(8,7,5,3,2,9,6,1,4)])
temp$Prop._h2 <- 100 * temp$Prop._h2
temp$lower <- temp$Prop._h2 - 196 * temp$Prop._h2_std_error
temp$upper <- temp$Prop._h2 + 196 * temp$Prop._h2_std_error
plot_h2(temp, "", xlab = '% of SNP heritability explained', ylab = "Functional annotations")
```

#### Dynamic elements vs. static elements from pseudo bulk (marginal analysis)
```{r, echo=F}
enrich.ld <- read.table("data/enrich_results/pseudobulk_dyn_vs_stat_marg_total.tsv", sep=",", header=T, row.names=1)
temp <- enrich.ld
temp$annotation <- factor(rownames(temp), level=rownames(temp)[c(1,2,5,4,8,3,9,6,7)])
temp$lower <- temp$Enrichment - 1.96 * temp$Enrichment_std_error
temp$upper <- temp$Enrichment + 1.96 * temp$Enrichment_std_error
plot_enrichment(temp, "", xlab = 'enrichment of SZ variants', ylab = "Functional annotations")
```





#### Diverse diseases (joint analysis)
```{r, echo=F}
enrich.ld <- read.table("data/enrich_results/various_ldsc_joint/intelligence_pseudobulk_dyn_vs_stat_joint.results", sep="\t", header=T, row.names=1)
temp <- enrich.ld[(nrow(enrich.ld)-8):nrow(enrich.ld),]
annot <- c('1vs0_NEFMp',
 '6vs0_GABA',
 '1vs0_GABA',
 '6vs0_NEFMp',
 'NpCm_glut_static',
 '6vs0_NEFMm',
 'NmCp_glut_static',
 'GABA_static',
 '1vs0_NEFMm')
rownames(temp) <- annot
temp$annotation <- factor(rownames(temp), level=rownames(temp)[c(8,7,5,3,2,9,6,1,4)])
temp$lower <- temp$Enrichment - 1.96 * temp$Enrichment_std_error
temp$upper <- temp$Enrichment + 1.96 * temp$Enrichment_std_error
plot_enrichment(temp, "", xlab = 'enrichment of SZ variants', ylab = "Functional annotations")
```


#### Enrichment analysis applied to various diseases
```{r, echo=F}
path <- "data/enrich_results/various_ldsc_joint/"
res.dir <- list.files(path, pattern="pseudobulk")
diseases <- sapply(strsplit(res.dir,"_"), function(x) {x[1]})
aggr <- list()
for(i in seq_along(res.dir)) {
  enrich.ld <- read.table(paste0(path, res.dir[i]), sep="\t", header=T, row.names=1)
  temp <- enrich.ld[(nrow(enrich.ld)-8):nrow(enrich.ld),]
  annot <- c('1vs0_NEFMp',
 '6vs0_GABA',
 '1vs0_GABA',
 '6vs0_NEFMp',
 'NpCm_glut_static',
 '6vs0_NEFMm',
 'NmCp_glut_static',
 'GABA_static',
 '1vs0_NEFMm')
  rownames(temp) <- annot
  aggr[[i]] <- data.frame(disease=diseases[i], annot=annot, enrichment=temp$Enrichment, enrichment_pval=-log10(temp$Enrichment_p))
}
aggr <- do.call("rbind", aggr)
aggr$enrichment[aggr$enrichment<0] = 0

aggr$annot <- factor(aggr$annot, level=annot[c(8,7,5,3,2,9,6,1,4)])
aggr$disease <- factor(aggr$disease, level=diseases[c(7,6,5,4,1,3,8,2)])

ggplot(data=aggr, aes(x=disease, y=annot, color=enrichment_pval, size=enrichment)) + 
  geom_point() +
  scale_color_gradient(low="black", high="red") +
  theme_bw() + 
  ylab("") + 
  xlab("") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(color="-log10(Pvalue)")
  ggtitle("S-LDSC enrichment analysis")
```





### Enrichment analysis for Signac peak sets
#### Dynamic elements vs. static elements (joint analysis)
We rerun stratified LD score regression with a smaller peak sets of resting condition (0hr). 0hr peaks are called through MACS2-signac. The number of peaks are much smaller. 0hr peaks are more enriched than the differential peaks. Still, when the differential elements were removed from 0hr peaks ("static" peaks in the plot), the enrichment is reduced dramatically. 
```{r, echo=F}
ld.smaller <- read.table("data/enrich_results/Signac_dyn_vs_stat_joint.results", sep="\t", header=T, row.names=1)
ld.smaller <- ld.smaller[(nrow(ld.smaller)-8):nrow(ld.smaller),]
rownames(ld.smaller) <- c('NmCp_glut_static',
 'NpCm_glut_static',
 'GABA_static',
 'GABA_1v0hr',
 'GABA_6v0hr',
 'NmCp_glut_1v0hr',
 'NpCm_glut_6v0hr',
 'NpCm_glut_1v0hr',
 'NmCp_glut_6v0hr')
ld.smaller$annotation <- factor(row.names(ld.smaller), levels=rev(row.names(ld.smaller)[c(1:6,9,8,7)]))
ld.smaller$lower <- ld.smaller$Enrichment - 1.96 * ld.smaller$Enrichment_std_error
ld.smaller$upper <- ld.smaller$Enrichment + 1.96 * ld.smaller$Enrichment_std_error

plot_enrichment(ld.smaller, "", xlab = 'enrichment of SZ variants', ylab = "Functional annotations")
```

#### Comparison of heritability of peak sets (joint analysis)
```{r, echo=F}
ld.smaller <- read.table("data/enrich_results/Signac_dyn_vs_stat_joint.results", sep="\t", header=T, row.names=1)
ld.smaller <- ld.smaller[(nrow(ld.smaller)-8):nrow(ld.smaller),]
rownames(ld.smaller) <- c('NmCp_glut_static',
 'NpCm_glut_static',
 'GABA_static',
 'GABA_1v0hr',
 'GABA_6v0hr',
 'NmCp_glut_1v0hr',
 'NpCm_glut_6v0hr',
 'NpCm_glut_1v0hr',
 'NmCp_glut_6v0hr')
ld.smaller$annotation <- factor(row.names(ld.smaller), levels=rev(row.names(ld.smaller)))
ld.smaller$Prop._h2 <- 100 * ld.smaller$Prop._h2
ld.smaller$lower <- ld.smaller$Prop._h2 - 196 * ld.smaller$Prop._h2_std_error
ld.smaller$upper <- ld.smaller$Prop._h2 + 196 * ld.smaller$Prop._h2_std_error

plot_h2(ld.smaller, "", xlab = '% of SNP heritability explained', ylab = "Functional annotations")
```


#### Dynamic elements vs. static elements (marginal analysis)
We run S-LDSC with each annotation one at a time. Baseline model was always included.
```{r, echo=F}
ld.smaller <- read.table("data/enrich_results/ldsc_marginal_dynamic_vs_static.tsv", sep=",", header=T, row.names=1)
ld.smaller <- ld.smaller[-grep("psbulk", rownames(ld.smaller)),]
ld.smaller <- ld.smaller[-grep("NPC", rownames(ld.smaller)),]
ld.smaller$annotation <- factor(row.names(ld.smaller), 
           levels=rownames(ld.smaller)[c(9,2,8,4,5,6,1,3,7)])
ld.smaller$lower <- ld.smaller$Enrichment - 1.96 * ld.smaller$Enrichment_std_error
ld.smaller$upper <- ld.smaller$Enrichment + 1.96 * ld.smaller$Enrichment_std_error

plot_enrichment(ld.smaller, "", xlab = 'enrichment of SZ variants', ylab = "Functional annotations")
```


### Applying to diverse diseases (joint analysis)

```{r, echo=F}
ld.smaller <- read.table("data/enrich_results/various_ldsc_joint/neuroticism_Signac_dyn_vs_stat_joint.results", sep="\t", header=T, row.names=1)
ld.smaller <- ld.smaller[(nrow(ld.smaller)-8):nrow(ld.smaller),]
rownames(ld.smaller) <- c('NmCp_glut_static',
 'NpCm_glut_static',
 'GABA_static',
 'GABA_1v0hr',
 'GABA_6v0hr',
 'NmCp_glut_1v0hr',
 'NpCm_glut_6v0hr',
 'NpCm_glut_1v0hr',
 'NmCp_glut_6v0hr')
ld.smaller$annotation <- factor(row.names(ld.smaller), levels=rev(row.names(ld.smaller)[c(1:6,9,8,7)]))
ld.smaller$lower <- ld.smaller$Enrichment - 1.96 * ld.smaller$Enrichment_std_error
ld.smaller$upper <- ld.smaller$Enrichment + 1.96 * ld.smaller$Enrichment_std_error

plot_enrichment(ld.smaller, "", xlab = 'enrichment of SZ variants', ylab = "Functional annotations")
```

#### Enrichment analysis applied to various diseases
```{r, echo=F}
path <- "data/enrich_results/various_ldsc_joint/"
res.dir <- list.files(path, pattern="Signac")
diseases <- sapply(strsplit(res.dir,"_"), function(x) {x[1]})
aggr <- list()
for(i in seq_along(res.dir)) {
  enrich.ld <- read.table(paste0(path, res.dir[i]), sep="\t", header=T, row.names=1)
  temp <- enrich.ld[(nrow(enrich.ld)-8):nrow(enrich.ld),]
  annot <- c('NmCp_glut_static',
 'NpCm_glut_static',
 'GABA_static',
 'GABA_1v0hr',
 'GABA_6v0hr',
 'NmCp_glut_1v0hr',
 'NpCm_glut_6v0hr',
 'NpCm_glut_1v0hr',
 'NmCp_glut_6v0hr')
  rownames(temp) <- annot
  aggr[[i]] <- data.frame(disease=diseases[i], annot=annot, enrichment=temp$Enrichment, enrichment_pval=-log10(temp$Enrichment_p))
}
aggr <- do.call("rbind", aggr)
aggr$enrichment[aggr$enrichment<0] = 0

aggr$annot <- factor(aggr$annot, level=annot[c(1:6,9,8,7)])
aggr$disease <- factor(aggr$disease, level=diseases[c(7,6,5,4,1,3,8,2)])

ggplot(data=aggr, aes(x=disease, y=annot, color=enrichment_pval, size=enrichment)) + 
  geom_point() +
  scale_color_gradient(low="black", high="red") +
  theme_bw() + 
  ylab("") + 
  xlab("") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(color="-log10(Pvalue)")
  ggtitle("S-LDSC enrichment analysis")
```
