---
title: "TORUS enrichemnt"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    number_sections: no
    toc: yes
    toc_float: true
editor_options:
  chunk_output_type: console
---

## TORUS enrichment analysis
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(rtracklayer)

plot_or = function(dat, title, xlab="Odds ratio", ylab="Annotations") {
  p1 <- ggplot(dat, aes(y=annotation)) + 
    geom_point(aes(x=logOR), colour="#800000") + #plot OR
    geom_segment(aes(x=lower,xend=upper,yend=annotation), alpha=0.8)
  p1 <- p1 + geom_vline(xintercept=0) + theme_bw() + labs(title=title, x=xlab, y=ylab)
  p1
}
```

### Enrichment of positive control

The baseline used in the last paper was adopted here, including promoter regions, enhancer regions, and conserved regions.

```{r positive_control, echo=FALSE}
enrich.pos <- read.table("data/enrich_results/positive_control.tsv", header=T)
enrich.pos[,2:4] <- enrich.pos[,2:4] * log2(exp(1))
plot_or(enrich.pos, "", xlab = expression(paste(log[2], ' fold enrichment of SZ variants')),
        ylab = "Functional annotations")
```

### Peaks called at specific cell type and time point
There are two approaches to call the peaks. One is called through MACS2-Signac (named "Signac" later for the source. The other is merging samples within the same cell types and the same time point and call peaks with MACS2 (called "pseudo bulk" later). The pseudo bulk peakset has at least twice as much peaks as the Signac peak set across all cell types and time points (refer to the [the peak counts](index.html#summary-for-different-peak-sets)).
```{r enrichment_12peaks, echo=FALSE}
enrich.12peaks <- read.table("data/enrich_results/12peaks.tsv", header=T)
enrich.12peaks[,2:4] <- enrich.12peaks[,2:4] * log2(exp(1))
plot_or(enrich.12peaks, "", xlab = expression(paste(log[2], ' fold enrichment of SZ variants')), ylab = "Functional annotations")
```

```{r enrichment_pseudo_bulk, echo=FALSE}
enrich.pseudo <- read.table("data/enrich_results/pseudobluk.tsv", header=T)
colnames(enrich.pseudo)  = c("annotation", "logOR", "lower", "upper")
# table <- rbind(enrich.pseudo, enrich.static)
table <- enrich.pseudo
table[,2:4] <- table[,2:4] * log2(exp(1))
table$annotation <- factor(table$annotation, levels=table$annotation)
plot_or(table[-c(5,6,11),], "", xlab = expression(paste(log[2], ' fold enrichment of SZ variants')), ylab = "Functional annotations")
```

### Enrichment analysis of FWER<0.25
The number of peaks selected with Bonferroni q values are too small even with 0.25 cutoff.
```{r, echo=F}
enrich.bonf25 <- read.table("data/enrich_results/bonf25_bothside.tsv", header=T)
colnames(enrich.bonf25)  = c("annotation", "logOR", "lower", "upper")
plot_or(enrich.bonf25, "", xlab = expression(paste(log[2], ' fold enrichment of SZ variants')), ylab = "Functional annotations")
```

### DAP at FDR<0.05
After recomputing FDR with BH procedure, differentially accessibility peaks (DAP) were selected at FDR < 0.05. The number of peaks are much larger than FWER peak sets (refer to [the peak counts](index.html#summary-for-different-peak-sets)).

```{r enrichment_DAP_bothside, echo=FALSE}
enrich.dap <- read.table("data/enrich_results/FDR5_bothside.tsv", header=T)
colnames(enrich.dap)  = c("annotation", "logOR", "lower", "upper")
enrich.dap[,2:4] <- enrich.dap[,2:4] * log2(exp(1))
plot_or(enrich.dap[-(5:6),], "", xlab = expression(paste(log[2], ' fold enrichment of SZ variants')), ylab = "Functional annotations")
```

<!-- ### Inducible elements -->
<!-- ```{r enrichment_DAP_positive, echo=FALSE} -->
<!-- enrich.dap.pos <- read.table("data/enrich_results/FDR5_positive.tsv", header=T) -->
<!-- colnames(enrich.dap.pos)  = c("annotation", "logOR", "lower", "upper") -->
<!-- plot_or(enrich.dap.pos[-(5:6),], "", xlab = expression(paste(log[2], ' fold enrichment of SZ variants')), ylab = "Functional annotations") -->
<!-- ``` -->

#### Dynamic vs. static elements from Signac (marginal analysis)
We remove peaks in the resting condition (0 hr) peak sets that overlap with the differential peaks from the corresponding cell type.

```{r enrichment_vs_signac, echo=FALSE}
enrich.dap <- read.table("data/enrich_results/FDR5_bothside.tsv", header=T)
colnames(enrich.dap)  = c("annotation", "logOR", "lower", "upper")
enrich.static <- read.table("data/enrich_results/static_bothside.tsv", header=T)
colnames(enrich.static)  = c("annotation", "logOR", "lower", "upper")
table <- rbind(enrich.dap, enrich.static)
table[,2:4] <- table[,2:4] * log2(exp(1))
plot_or(table[-c(5,6,11),], "", xlab = expression(paste(log[2], ' fold enrichment of SZ variants')), ylab = "Functional annotations")
```

#### Dynamic vs. static elements from Signac (joint analysis)
```{r enrichment_vs_signac_joint, echo=FALSE}
enrich.joint <- read.table("data/joint_dynamic/joint_dynamic_vs_static.tsv", header=T)
colnames(enrich.joint)  = c("annotation", "logOR", "lower", "upper")
table <- enrich.joint[-c(1,4,10,11),]
table[,2:4] <- table[,2:4] * log2(exp(1))
table$annotation <- factor(table$annotation, levels=table$annotation)
plot_or(table, "", xlab = expression(paste(log[2], ' fold enrichment of SZ variants')), ylab = "Functional annotations")
```


#### Dynamic vs. static elements from pseudo bulk (marginal analysis)
```{r enrichment_vs_pseudobulk, echo=FALSE}
enrich.dap <- read.table("data/enrich_results/marginal_pseudobulk_dynamic_vs_static.tsv", header=T)
colnames(enrich.dap)  = c("annotation", "logOR", "lower", "upper")
# enrich.static.pseu <- read.table("data/enrich_results/static_bothside.tsv", header=T)
# colnames(enrich.static.pseu)  = c("annotation", "logOR", "lower", "upper")
table <- enrich.dap[-c(3,8,12),]
table[,2:4] <- table[,2:4] * log2(exp(1))
table$annotation <- factor(table$annotation, levels=table$annotation)
plot_or(table, "", xlab = expression(paste(log[2], ' fold enrichment of SZ variants')), ylab = "Functional annotations")
```

#### Dynamic vs. static elements from pseudo bulk (joint analysis)
Comparison with pseudo bulk in joint analysis show different results. This is probably because the differential peaks were not derived from the pseudo bulk approach. Additional peaks gained from this approach have not been eliminated.

```{r enrichment_vs_pseudobulk_joint, echo=FALSE}
enrich.joint.pseu <- read.table("data/enrich_results/joint_pseudobulk_dynvs_stat.tsv", header=T)
colnames(enrich.joint.pseu)  = c("annotation", "logOR", "lower", "upper")
table <- enrich.joint.pseu[-c(1),]
table[,2:4] <- table[,2:4] * log2(exp(1))
table$annotation <- factor(table$annotation, levels=table$annotation)
plot_or(table, "", xlab = expression(paste(log[2], ' fold enrichment of SZ variants')), ylab = "Functional annotations")
```